---
title: "Creating Data"
author: "Benjamin Egan"
output: html_document
---

#### Created 4/3/25

```{r setup, include = FALSE}

library(tidyverse) 
library(ggplot2)

```

## Objective

As I mentioned in piece 3, I serve as a teaching assistant for the undergraduate Research Methods in psychology course at Wake Forest University. They are required to write a final paper, for which they need to write up and analyze a 2x2 experiment. As a TA, I have to create the data for the experiment. Since it's a 2x2 between subjects with at least 100 participants, this would be incredibly difficult. Luckily I walked through module 13 on simulating data, and should be able to create the data.

The basic function I'll need to use is rnorm(). This will let me randomly generate a sample of a normal distribution. I'm starting and making sure I can actually generate the data.


```{r base rnorm}

set.seed(123)

base_rnorm <- rnorm(100, mean = 50, sd = 5)

first_attempt <- data.frame(id = 1:100, base_rnorm)

                                      
data.frame(first_attempt) %>%
ggplot(
 aes(x=base_rnorm)
)+
  theme_bw()+
  geom_histogram(binwidth=2,color="black", fill="blue") +
  labs(
    title = "Look, I can create data!",
    x = NULL, 
    y = NULL
    )

```

My plan is to create four different samples, one for each group. I then will join them to create a large dataset. This way I can control the main effects and interactions.

The experiment calls for two variables, Sugar (sugar/no sugar) and Caffeine (caffeine/no caffeine), and the DV is reported distance to an object.

## Creating the groups

### Sugar and caffeine
```{r sugar and caffeine}

set.seed(101)

Distance <- rnorm(30, mean = 100, sd = 4.03)

group1 <- data.frame(id = 1:30, Sugar = 1, Caffeine = 1, Distance)


```

### Sugar and no caffeine
```{r sugar and no caffeine}

set.seed(102)

Distance <- rnorm(30, mean = 135, sd = 5.1)

group2 <- data.frame(id = 31:60, Sugar = 1, Caffeine = 0, Distance)

```

### No sugar and caffeine
```{r no sugar and caffeine}

set.seed(103)

Distance <- rnorm(30, mean = 120, sd = 4.3)

group3 <- data.frame(id = 61:90, Sugar = 0, Caffeine = 1, Distance)


```


### No sugar and no caffeine
```{r neither sugar and caffeine}

set.seed(104)

Distance <- rnorm(30, mean = 160, sd = 6.05)

group4 <- data.frame(id = 91:120, Sugar = 0, Caffeine = 0, Distance)

```

## Joining the groups

```{r join groups}

group1and2 <- group1 %>% full_join(group2)

group3and4 <- group3 %>% full_join(group4)

lab_12_data <- group1and2 %>% full_join(group3and4)

write_csv(lab_12_data, "testdata.csv")

hello <- read.csv("testdata.csv")

```

lab_12_data is a data frame that includes an ID number, if they had sugar or caffeine, and a distance score.

```{r anova test}



```

